Providing a Unified Software Environment for Canada’s National Advanced Computing Centers

### 2019-04-09 paper.tex
Merged in Bart's changes.

%
% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title{Providing a Unified Software Environment for Canada’s National Advanced Computing Centers}

@@

\begin{figure*}
  \includegraphics[width=0.7\textwidth]{layers.png}
  \caption{Layered structure of our software stack. The bottom layer (red) is entirely managed by the system administrators of each cluster. The second layer (grey) contains libraries that can be installed centrally, through Nix, but can be overridden locally. The third layer (yellow) contains everything that one would typically find in the Linux user space. The fourth layer (green) contains packages that you would typically find in an operating system, but for which multiple versions are required. They are installed through Nix, but wrapped into a module by EasyBuild. The top layer (blue) contains packages that are not typically installed in a Linux distribution, and are installed via EasyBuild.}
  \label{tab:layers}
\end{figure*}

@@

@@

If you plan to use our CVMFS repositories on a large number of nodes such as a cluster, we require that you deploy one or more caching HTTP proxies, to avoid overloading the stratum 1 servers. If you are interested in testing or making our software environment available to your users on your cluster, we ask that you first contact us as described in the webpage mentioned before. Note that this software stack comes with absolutely no warranty.

>>>>

%
% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{colortbl}
\usepackage{amssymb}
%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title[A Unified Software Environment for Canada’s Advanced Computing Centers]{Providing a Unified Software Environment for Canada’s National Advanced Computing Centers}

@@

\begin{figure*}
  \begin{center}
  {\Large
  \begin{tabular}{|p{0.7\textwidth}|}
  \hline \cellcolor{cyan}
  EasyBuild layer: modules for Intel, PGI, Open MPI, MKL, high-level applications.
  Multiple instruction sets (SSE3, AVX, AVX2, AVX512).\newline
  \texttt{/cvmfs/soft.computecanada.ca/easybuild/\{modules,software\}/2017}\\
  \hline \cellcolor{green}
  EasyBuild-generated modules around Nix profiles: for example Clang, Eclipse,
  Erlang, FPC, GCC, Ruby.\newline
  \texttt{/cvmfs/soft.computecanada.ca/nix/var/nix/profiles/[a-z]\**}\\
  \hline \cellcolor{yellow}
  Nix layer: GNU C Library (glibc), autotools, make, bash, cat, ls, awk, grep, etc. \newline
  \texttt{module load nixpkgs/16.09 $\Rightarrow$ \$EBROOTNIXPKGS=} \newline
  \texttt{/cvmfs/soft.computecanada.ca/nix/var/nix/profiles/16.09}\\
  \hline \cellcolor{lightgray} 
  Gray area: Lustre client libraries, IB/OmniPath/InfiniPath client libraries 
  (all dependencies of Open MPI). In Nix layer, but can be overridden using
  \texttt{PATH} and \texttt{LD\_LIBRARY\_PATH}.\\
  \hline \cellcolor{pink}
  OS kernel, daemons, drivers, libcuda, anything privileged (e.g. the sudo
  command): always local. Some legally restricted software too, such as VASP.\\
  \hline
  \end{tabular}
  }
  \end{center}
  \caption{Layered structure of our software stack. The bottom layer (pink) is entirely managed by the system administrators of each cluster. The second layer (grey) contains libraries that can be installed centrally, through Nix, but can be overridden locally. The third layer (yellow) contains everything that one would typically find in the Linux user space. The fourth layer (green) contains packages that you would typically find in an operating system, but for which multiple versions are required. They are installed through Nix, but wrapped into a module by EasyBuild. The top layer (cyan) contains packages that are not typically installed in a Linux distribution, and are installed via EasyBuild.}
  \label{tab:layers}
\end{figure*}

@@

In general, Nix provides an environment that is internally consistent but is best developed using Nix itself, through tools such as \texttt{nix-shell}. Such tools can only be used if the Nix store is writable, an impossibility with CVMFS. Attempting to use Nix in conjunction with other development tools using environment variables can be fragile at times. Moreover, the dependency of the hash on the whole installation recipe, instead of only the dependencies (as, for example, implemented in Spack~\cite{Spack}), causes the store path to change even if a bug fix or security update needs to be applied, aggravating this problem. We must note however, that some clusters in the Compute Canada provide Nix with a writable Nix store, which is completely separate from the software stack we are describing here.

@@

If you plan to use our CVMFS repositories on a large number of nodes such as a cluster, we require that you deploy one or more caching HTTP proxies, to avoid overloading the stratum 1 servers. If you are interested in testing or making our software environment available to your users on your cluster, we ask that you first contact us as described in the web page mentioned before. Note that this software stack comes with absolutely no warranty.
